---
title: "IL Default Switch V2"
author: "Bryan Butler"
date: "May 17, 2019"
output:
    html_document:
    toc: false
    toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, cache.lazy = FALSE)
```

# <strong>IL Default Paperless Setting Impact</strong> {.tabset .tabset-fade .tabset-pills}


<style>
  .main-container {
    max-width: 1200px !important;
    margin-left: auto;
    margin-right: auto;
  }
</style>


## Overview of Analysis
### - Utilise IL Pre-Registration and Enrollments
### - Model the past up to February; generates existing run rate
### - Forecast March and April using existing run rate
### - Compare to actual change after the default setting changed


## Key Findings
### - Despite limited data able to make reasonable models of IL Pre-Registrations and Enrollments
#### - Error rate of fit is 16 - 17%
### - Using the model, we projected out the March and April values as if no change to the default setting occurred
### - For both series, the change in default significantly altered the March and April counts



```{r, loadPython}
library(reticulate)
use_condaenv('timeseries')

```



## Load Python
```{python}
import PyQt5
import pandas as pd
from pandas.plotting import lag_plot


import numpy as np
import os

import matplotlib.pyplot as plt
from matplotlib.pyplot import figure

import statsmodels.api as sm
from statsmodels.graphics.tsaplots import plot_acf, plot_pacf
from statsmodels.tsa.arima_model import ARMA, ARIMA, ARMAResults, ARIMAResults

from statsmodels.tsa.statespace.sarimax import SARIMAX

from statsmodels.tsa.seasonal import seasonal_decompose
from statsmodels.tsa.statespace.tools import diff

from statsmodels.tools.eval_measures import rmse

# ignore harmless warnings
import warnings
warnings.filterwarnings('ignore')

```


## Dickey Fuller Function
```{python, echo = T}
# get the Dickey-Fuller function


from statsmodels.tsa.stattools import adfuller

def adf_test(series,title=''):
    """
    Pass in a time series and an optional title, returns an ADF report
    """
    print(f'Augmented Dickey-Fuller Test: {title}')
    result = adfuller(series.dropna(),autolag='AIC') # .dropna() handles differenced data
    
    labels = ['ADF test statistic','p-value','# lags used','# observations']
    out = pd.Series(result[0:4],index=labels)

    for key,val in result[4].items():
        out[f'critical value ({key})']=val
        
    print(out.to_string())          # .to_string() removes the line "dtype: float64"
    
    if result[1] <= 0.05:
        print("Strong evidence against the null hypothesis")
        print("Reject the null hypothesis")
        print("Data has no unit root and is stationary")
    else:
        print("Weak evidence against the null hypothesis")
        print("Fail to reject the null hypothesis")
        print("Data has a unit root and is non-stationary")
```



## Get the Data
```{python}
os.chdir('/Users/Bryan/Documents/Programming/useR')
os.getcwd()
```


```{python}
# load the data create the index for dates

df = pd.read_csv('IL_Registrations.csv', index_col='Date', parse_dates=True)
df.index = pd.to_datetime(df.index, dayfirst=True)

# set up for monthly
df.index.freq = 'MS'

df.head()
```



## Plot the Series
```{python plotseries}
# view all of the series
ax = df.plot(figsize=(16,9), legend=True, title='Pre Registrations and Enrollments')
plt.show()
```


### adf test
```{python adf1}
adf_test(df['IL_Pre_Reg'])
```


## Reduce the Series and Plot
```{python plotreduced}
dfRed = df.iloc[1:12].copy()
ax = dfRed.plot(figsize=(16,9), legend=True, title='Plot of Reduced Data Set')
plt.show()
```


## Separate IL Data Only
```{python il}
# separate the IL data
IL = df[['IL_Pre_Reg', 'IL_Enroll']]


# remove the ends
ILred = IL.iloc[1:12]

print(ILred.head())

ax = ILred.plot(figsize=(16,6), legend=True)
plt.show()
```


### ADF Test on Pre-Registrations
```{python}
adf_test(ILred['IL_Pre_Reg'])
```


### ADF Test on Enrollments
```{python}
# adf test
adf_test(ILred['IL_Enroll'])
```


## Correlations - ACF and PACF
### Pre-Registrations
```{python, echo=F}
# check the ACF of Pre-Registrations
N, M = 12, 6
fig, ax = plt.subplots(figsize=(N, M))
plot_acf(ILred['IL_Pre_Reg'], lags=10, ax=ax)
plt.show()


# check PACF of Pre Registsrations

N, M = 12, 6
fig, ax = plt.subplots(figsize=(N, M))
plot_pacf(ILred['IL_Pre_Reg'], lags=10, ax=ax)
plt.show()
```


### Enrollments
```{python, echo=F}
# check the ACF of Pre-Enrollments

N, M = 12, 6
fig, ax = plt.subplots(figsize=(N, M))
plot_acf(ILred['IL_Enroll'], lags=10, ax=ax)
plt.show()


# check the PACF of Pre-Enrollments

N, M = 12, 6
fig, ax = plt.subplots(figsize=(N, M))
plot_pacf(ILred['IL_Enroll'], lags=10, ax=ax)
plt.show()


```




## Model PreRegistrations
```{python}
# split pre-reg

train = ILred['IL_Pre_Reg'].iloc[0:9]
test = ILred['IL_Pre_Reg'].iloc[9:]
```


```{python}
# select 0,1,1 for order

model = ARIMA(train, order=(0,1,1))
results = model.fit()
print(results.summary())

start = len(train)
end = len(train) + len(test) - 1

predictions = results.predict(start=start, end=end, typ='levels').rename('Modeled Run_Rate')


backtest = results.predict(start=1, end=start-1, typ='levels').rename('Modeled Run_Rate')
```


### Check Error
```{python}
error = rmse(train[1:], backtest)
print(error)

percent = error/train.mean()
print(percent)
```


### Comparison of Modeled, Actual and Forecasted for Pre-Registrations
```{python}
reg = ILred['IL_Pre_Reg'].copy()
dfreg = pd.concat([reg, backtest], axis = 1)

ax = dfreg.plot(figsize=(16,8), legend = True)
ax1 = predictions.plot(legend=True)
plt.show()


```


## Enrollments
```{python}
# split enrollements
train = ILred['IL_Enroll'].iloc[0:9]
test = ILred['IL_Enroll'].iloc[9:]

# select 3,0,0 for order

model = ARIMA(train, order=(3,0,0))
results = model.fit()
print(results.summary())


start = len(train)
end = len(train) + len(test) - 1

predictions = results.predict(start=start, end=end).rename('Enrollment Run Rate')

error = rmse(train[1:9], backtest)
print(error)

percent = error/train.mean()
print(percent)


backtest = results.predict(start=1, end=start-1).rename('Modeled Enrollments')
```


### Comparison of Modeled, Actual and Forecasted for Enrollments
```{python}
# make a plot of model fit
act = ILred['IL_Enroll'].copy()
dfenroll = pd.concat([act, backtest], axis = 1)

ax = dfenroll.plot(figsize=(16,8), legend = True)
ax1 = predictions.plot(legend=True)
plt.show()
```





