---
title: "Cancellations TS"
author: "Bryan Butler"
date: "7/1/2019"
output:
    html_document:
    toc: false
    toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, cache.lazy = FALSE)
```

# <strong>Time Series of Auto Policies</strong> {.tabset .tabset-fade .tabset-pills}


<style>
  .main-container {
    max-width: 1200px !important;
    margin-left: auto;
    margin-right: auto;
  }
</style>


## Overview of Analysis
### - Collect data and features on auto polcies (effective starts and cancellations)
### - Plot and model the series
### - Use ARIMA/SARIMAX models to backtest and forecast
### - Try LSTM (neural network model)


## Key Findings
### - TBD


```{r, loadPython}
library(reticulate)
use_condaenv('timeseries')

```


## Load Python
```{python importLibaries}

# Base libraries
import PyQt5
import pandas as pd
from pandas import Series, DataFrame
from pandas.plotting import lag_plot


import numpy as np
import pyodbc


import os
import math
from itertools import cycle


# plotting libraries
import matplotlib.pyplot as plt
import seaborn as sns


# get the datetime library for date & time calcs
from datetime import datetime, timedelta


# to visualize all of the columns in the dataframe
pd.pandas.set_option('display.max_columns', None)


from matplotlib.pyplot import figure



# stats models tools
import statsmodels.api as sm


# Dickey-Fuller test for stationarity
from statsmodels.tsa.stattools import adfuller


# seaonal decomposition
from statsmodels.tsa.seasonal import seasonal_decompose


# for expanding plots
from pylab import rcParams


# creating difference sequences
from statsmodels.tsa.statespace.tools import diff


# correlation plots
from statsmodels.graphics.tsaplots import plot_acf, plot_pacf


# modeling tools
from statsmodels.tsa.arima_model import ARMA, ARIMA, ARMAResults, ARIMAResults
from statsmodels.tsa.statespace.sarimax import SARIMAX


# metrics
from statsmodels.tools.eval_measures import mse, rmse


# ignore harmless warnings
import warnings
warnings.filterwarnings('ignore')
```



## Dickey Fuller Function
```{python adfTest, echo = T}
# get the Dickey-Fuller function


def adf_test(series,title=''):
    """
    Pass in a time series and an optional title, returns an ADF report
    """
    print(f'Augmented Dickey-Fuller Test: {title}')
    result = adfuller(series.dropna(),autolag='AIC') # .dropna() handles differenced data
    
    labels = ['ADF test statistic','p-value','# lags used','# observations']
    out = pd.Series(result[0:4],index=labels)

    for key,val in result[4].items():
        out[f'critical value ({key})']=val
        
    print(out.to_string())          # .to_string() removes the line "dtype: float64"
    
    if result[1] <= 0.05:
        print("Strong evidence against the null hypothesis")
        print("Reject the null hypothesis")
        print("Data has no unit root and is stationary")
    else:
        print("Weak evidence against the null hypothesis")
        print("Fail to reject the null hypothesis")
        print("Data has a unit root and is non-stationary")
```


## Get the Data
```{python getData}

# create the path for the file

myPath = os.path.join('N:', '\Bryan', 'Retention')

print(myPath)

os.chdir(myPath)

data = pd.read_csv('CombinedCancelNB.csv',index_col='StartDate', parse_dates=True)

# slice the series

print(f'{len(data)}', "rows in data")

print(data.head())

dataRed = data.iloc[59:107].copy()

print(dataRed.head())

print(f'{len(dataRed)}', "rows in dataRed")

```



## Plot the Series
```{python plotseries}
# view all of the series
data.plot(figsize=(16,9), legend=True, title = 'Relationship Between Cancellations and New Customers for Auto')
plt.show()
```


```{python plotseriesblowup}
# now blow it up to scale

title = 'Relationship Between Cancellations and New Auto Customers'
ylabel = 'Counts per Month'
xlabel = ''

# ax = dataRed['NewBus'].plot(legend=True, figsize=(16,6),title=title)
# ax1 = dataRed['Cancellations'].plot(legend=True)
# ax.autoscale(axis='x', tight=True)
# ax.set(xlabel=xlabel, ylabel=ylabel)
# plt.show()

dataRed.plot(figsize=(16,9), legend=True, title=title)
plt.show()



# get a view 
print(dataRed['Cancellations'].head(15))
```


